### 2019-09-25 reports/thesis/chapters/4_variancePrediction.tex
Revise and used proofreader
Final revise

The aim is to model and predict the grey value variance of a pixel given the its grey value by using a model fitted onto the sample variance-mean data. No spatial information was used, thus, the variance-mean data is denoted as $(x_1,y_1),(x_2,y_2),\dotsc,(x_N,y_N)$ from now on, where $N$ is the area of the ROI or the size of the dataset. Let $Y(x)$ be a random variable and the sample variance given a grey value mean $x$. It was assumed that the standard error from estimating the mean was negligible so that the uncertainty is captured by the random variable $Y$. It was assumed that for a given pixel, the grey values are Normal and i.i.d. Let $\sigma^2(x)$ be the variance given a grey value $x$, then it can be shown that
\begin{equation}
\dfrac{(m-1)Y(x)}{\sigma^2(x)}\sim\chi^2_{m-1}
\end{equation}
which results in
\begin{equation}
Y(x)\sim\gammaDist\left(\alpha,\dfrac{\alpha}{\sigma^2(x)}\right)
\end{equation}
where $\alpha=(m-1)/2$ is the shape parameter. The expectation and variance are
\begin{equation}
\expectation\left[Y(x)\right]=\sigma^2(x)
\end{equation}
and
\begin{equation}
\variance\left[Y(x)\right]=\left(\dfrac{\sigma^2(x)}{\alpha}\right)^2
\end{equation}
respectively.

@@

The mean scaled deviance was used to assess the performance of variance prediction of the test set. The deviance is defined to be
\begin{equation}
  D = 2\left(
    \ln L_\mathrm{s} - \ln L_\tau
  \right)
\end{equation}
where
\begin{equation}
\ln L_\tau = \sum_{i\in\tau} \left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\alpha\ln \widehat{y}_i
    +(\alpha-1)\ln y_i
    -\frac{\alpha y_i}{\widehat{y}_i}
  \right]
\end{equation}
is the log-likelihood of the test set given the fitted model and $\ln L_\mathrm{s}$ is the saturated likelihood. This is obtained by replacing all $\widehat{y}_i$ with $y_i$ in $\ln L_\tau$ so that
\begin{equation}
  \ln L_\mathrm{s} = \sum_{i\in\tau}\left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\ln y_i
    -\alpha
  \right]
  \ .
\end{equation}
Following from this, the deviance is
\begin{equation}
  D = 2\alpha
  \sum_{i\in\tau}\left[
      \dfrac{
          y_i-\widehat{y}_i
      }
      {
          \widehat{y}_i
      }
      - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
  \right]
  \ .
\end{equation}
The mean scaled deviance is obtained by removing the factor of $\alpha$ and dividing by $N_\tau$ to get
\begin{equation}
    D_\mathrm{s} = \dfrac{2}{N_\tau}
    \sum_{i\in\tau}\left[
        \dfrac{
            y_i-\widehat{y}_i
        }
        {
            \widehat{y}_i
        }
        - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
    \right]
    \ .
\end{equation}

\begin{figure}
  \centering
  \includegraphics[width=\subSize]{../figures/varmean/devianceGraph.eps}
  \caption{Scaled deviance loss function}
  \label{fig:meanVar_deviance}
\end{figure}

@@

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsNoFilterDeg30_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsNoFilter} at \ang{30}}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsNoFilterDeg120_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsNoFilter} at \ang{120}}
  \end{subfigure}
  }
    \centerline{
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsFilterDeg30_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsFilter} at \ang{30}}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsFilterDeg120_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsFilter} at \ang{120}}
  \end{subfigure}
  }
  \caption{Test mean scaled deviance from predicting variances in a test set using a GLM fitted onto a training set. The different colours represent different shading corrections. The boxplot represent the 100 repeats of the analysis by bootstrapping the replicate projections.}
  \label{fig:meanVar_varMeanCv}
\end{figure}

>>>>

The aim is to model and predict the grey value variance of a pixel given its grey value by using a model fitted onto the sample variance-mean data. No spatial information was used, thus, the variance-mean data is denoted as $(x_1,y_1),(x_2,y_2),\dotsc,(x_N,y_N)$ from now on, where $N$ is the area of the ROI or the size of the dataset. Let $Y(x)$ be a random variable and the sample variance given a grey value mean $x$. It was assumed that the standard error from estimating the mean was negligible so that the uncertainty is captured by the random variable $Y$. It was assumed that for a given pixel, the grey values are Normal and i.i.d. Let $\sigma^2(x)$ be the variance given a grey value $x$, then it can be shown that
\begin{equation}
\dfrac{(m-1)Y(x)}{\sigma^2(x)}\sim\chi^2_{m-1}
\end{equation}
which results in
\begin{equation}
Y(x)\sim\gammaDist\left(\alpha,\dfrac{\alpha}{\sigma^2(x)}\right)
\end{equation}
where $\alpha=(m-1)/2$ is the shape parameter. The expectation and variance are
\begin{equation}
\expectation\left[Y(x)\right]=\sigma^2(x)
\end{equation}
and
\begin{equation}
\variance\left[Y(x)\right]=\left(\dfrac{\sigma^2(x)}{\alpha}\right)^2
\end{equation}
respectively.

@@

The mean scaled deviance was used to assess the performance of variance prediction of the test set. The deviance is defined to be
\begin{equation}
  D = 2\left(
    \ln L_\mathrm{s} - \ln L_\tau
  \right)
\end{equation}
where
\begin{equation}
\ln L_\tau = \sum_{i\in\tau} \left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\alpha\ln \widehat{y}_i
    +(\alpha-1)\ln y_i
    -\frac{\alpha y_i}{\widehat{y}_i}
  \right]
\end{equation}
is the log-likelihood of the test set given the fitted model and $\ln L_\mathrm{s}$ is the saturated log-likelihood. This is obtained by replacing all $\widehat{y}_i$ with $y_i$ in $\ln L_\tau$ so that
\begin{equation}
  \ln L_\mathrm{s} = \sum_{i\in\tau}\left[
    \alpha\ln\alpha
    -\ln\Gamma(\alpha)
    -\ln y_i
    -\alpha
  \right]
  \ .
\end{equation}
Following from this, the deviance is
\begin{equation}
  D = 2\alpha
  \sum_{i\in\tau}\left[
      \dfrac{
          y_i-\widehat{y}_i
      }
      {
          \widehat{y}_i
      }
      - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
  \right]
  \ .
\end{equation}
The mean scaled deviance is obtained by removing the factor of $\alpha$ and dividing by $N_\tau$ to get
\begin{equation}
    D_\mathrm{s} = \dfrac{2}{N_\tau}
    \sum_{i\in\tau}\left[
        \dfrac{
            y_i-\widehat{y}_i
        }
        {
            \widehat{y}_i
        }
        - \ln\left(\dfrac{y_i}{\widehat{y}_i}\right)
    \right]
    \ .
\end{equation}

\begin{figure}
  \centering
  \includegraphics[width=\subSize]{../figures/varmean/devianceGraph.eps}
  \caption{The scaled deviance loss function}
  \label{fig:meanVar_deviance}
\end{figure}

@@

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsNoFilterDeg30_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsNoFilter} at \ang{30}}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsNoFilterDeg120_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsNoFilter} at \ang{120}}
  \end{subfigure}
  }
    \centerline{
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsFilterDeg30_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsFilter} at \ang{30}}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
      \includegraphics[width=\textwidth]{../figures/varmean/VarMeanCvAbsFilterDeg120_testmeanscaleddeviance.eps}
      \caption{\texttt{AbsFilter} at \ang{120}}
  \end{subfigure}
  }
  \caption{Test mean scaled deviance from predicting variances in a test set using a GLM fitted onto a training set. The different colours represent different shading corrections. The boxplots represent the 100 repeats of the analysis by bootstrapping the replicate projections.}
  \label{fig:meanVar_varMeanCv}
\end{figure}
